Quick, Draw! Image Classification

### 2018-12-07 report/draft.tex
updating results section

%
% DS 5220 Final
%
\documentclass[12pt]{article}

@@

\subsubsection{$H2_a$ CNN should perform better than fully connected neural
  network}

@@

@@

In order to perform this experiment we built a fully connected neural network
using tensorflow. Our train:test split was chosen to be 80:20, an industry
standard.

Our initial objective was to build a model that overfit our given dataset,
with the idea that once our model is complex enough to overfit the training
data, we could regularize it in order to achieve a more optimal fit.

After experimenting with various architectures, we narrowed down to a 3 layer
model as specified in Figure \ref{fig:fcnn}.

The training set accuracy observed was 93\%, while the test accuracy was at
73\%, which clearly indicated an overfit in the model, as designed.

The model took 1500 epochs to converge taking around ~45 minutes on a
machine with 13 GB of RAM, Intel Xeon CPU 2.3 GHz (1 core, 2 threads),
1xTesla K80 GPU having 2496 CUDA cores with 12 GB GDDR5 VRAM. In order to
manage variance, we attempted regularizing using dropouts. Unfortunately,
that effort was futile. We then doubled the amount of data sampled to train
the model, eventually helping us achieve training set accuracy of 86\% and
test set accuracy of 82\%. Our learning rate was specified as 0.001.

@@

As for the CNN, we started off with the same approach with attempting to
overfit the model using convolution operations and a fully connected layer.
Since our objective is to categorise 10 categories, softmax was chosen as the
activation function for the final layer, while every other neuron had ReLU as
the activation function owing to ReLU’s faster convergence.

Number of Trainable Parameters: 320,890

The number of clearly being much greater than the number of observations used
to train the model, we need to reduce the number of parameters vis-a-vis
number of observations. From the available strategies to do so, we
incorporated MaxPooling [TODO: why?] post each convolution operation, in
order to extract only the most significant features. As as a consequence,
this reduced the volume of the output, giving the model a smaller memory
footprint as well as a smaller number of trainable parameters.

The observable training accuracy is 0.90, whereas the validation accuracy is
0.89. The model converged after 10 epochs, which took ~ 3 minutes.

Some of the other unsuccessful strategies used to decrease the number of
parameters and to regularize the network were strides and dropouts which
sacrificed way too much in the way of accuracy and thus had to be dropped.
Since most of our drawings had most of their information in the center of the
image, padding, in theory wouldn’t be of much help. However, actually
implementing padding improved our accuracy by 1\% at a cost of ~6000
parameters.

Conclusion

@@

ok

@@

@@

>>>>

%
% DS 5220 Final
%
\documentclass[11pt]{article}

@@

\subsubsection{$H2_a$: CNN should perform better than fully connected neural
  network}

@@

\subsubsection{$H3_a$: Small amounts of data will lead to overfitting}

Once we have found an optimal model for the dataset, we use this model
to study the effect of reduction in the sampled data. The optimal model
helps inform the relative performance of models trained with lower amounts
of data.

\subsubsection{$H4_a$: Imbalanced class prediction can be improved using data
  augmentation}

We tried to show the importance of having equal partitions of data to train
a machine learning model. Imbalanced classes can make a model biased toward
the majority class. We trained the neural network twice, on an imbalanced
dataset, keeping the number of samples for the majority class (fork), fixed
and decreasing the number of samples for the minority class (hammer).

The dataset was split into 80\% and 20\% for train and test respectively.
This split had 0.2\%, 2\%, 10\%, 20\%, 50\%, and 100\% of majority class
observations for the minority class. We designed the test dataset with 20000
fresh samples from the main dataset to get a clearer picture of the model
performance and we found that the accuracy on this fresh test dataset kept
on increasing as the number of samples in the minority class increased.

\subsubsection{$H5_a$ \& $H6_a$: Addition of a category to the trained model will reduce the performance of the model \& Adding more training data will improve the performance of the model}

We try and understand the limits and capacity of the finalized model by
varying the number of categories and the number of rows per category. We
then measure the response accuracy and response loss of the model. The
number of rows per category is altered in steps of 1000 from 7000 to
20000 and the number of categories is altered in steps of 1 from 3 to 10
categories. We attempt to formalize the patterns we see from visualizing
the data with an OLS regression.

@@

\subsubsection{$H1_a$\: Fully Connected Neural Network will struggle to
  get reasonable accuracy}

@@

\subsubsection{$H2_a$\: CNN should perform better than fully connected neural
  network}

@@

\subsubsection{$H3_a$\: Small amounts of data will lead to overfitting}

As found in the experiments, after decreasing the amount of data resulted
in overfitting thereby confirming our hypothesis.

\subsubsection{$H4_a$\: Imbalanced class prediction can be improved using
  data augmentation}

We get unexpectedly high accuracy of 43\% even in the worst case of 1:500
imbalance in the categories while attempting to classify only the minority
categories. This is because neural networks tend to generalize well on the
most important weights.

\subsubsection{$H5_a$ \& $H6_a$\: Addition of a category to the trained model
  will reduce the performance of the model/Adding more training data will
  improve the performance of the model}

It can be interpreted from the regression that for every additional category,
we can expect the test accuracy go down by 1.88\% and test loss to go up by
6.4\% while for every additional datapoint, we can expect the accuracy to go
up by .0016\% and loss to go down by .0056\%. Both results are statistically
significant with reasonably tight confidence intervals and standard errors.

@@

% Stats summary figures

@@

% Fully Connected Neural Network specs

% CNN model specs

% H3 results

\begin{figure}[h!]
  \begin{center}
    \includegraphics[scale=0.5]{fig8}
    \end{center}
  \caption{$H3_a$ Results for various size training data}
  \label{fig:h3Results}
  \end{figure}

% H4 results

\begin{figure}[h!]
  \begin{center}
    \includegraphics[scale=0.5]{fig9}
    \end{center}
  \caption{$H4_a$ Results for various class imbalances}
  \label{fig:h4Results}
\end{figure}

% H5 results

\begin{figure}[h!]
  \begin{center}
    \includegraphics[scale=0.5]{fig10}
    \end{center}
  \caption{$H5_a$ Results for various number of classes $k$}
  \label{fig:h5Results}
\end{figure}

% H6 results

\begin{figure}[h!]
  \begin{center}
    \includegraphics[scale=0.5]{fig11}
    \end{center}
  \caption{$H6_a$ Results for various amounts of training data $n$}
  \label{fig:h6Results}
  \end{figure}
